### 2015-05-13 paper/paper.tex
Added case study overview content

\documentclass[a4paper,11pt]{article}

\usepackage[paper=a4paper,left=30mm,width=150mm,top=25mm,bottom=25mm]{geometry}
\usepackage[hidelinks]{hyperref}    % All references will auto hyperlink
\usepackage[font={small}]{caption}  % Make figure captions smaller
\usepackage{graphicx}               % For the displaying of images in floats
\usepackage{mathtools}              % For more advanced math usage
\usepackage{poltakmacros}           % Personal macro package (https://gist.github.com/poltak/6ed59e8f30c9c399143f)

\begin{document}

\include{titlepage}                 % Pull in title page

\newpage
\pagenumbering{roman}
\tableofcontents
\newpage

\pagenumbering{arabic}

\section{Introduction} % (fold)
\label{sec:introduction}

While relational database management systems (RDBMS) have been somewhat of a ``go-to'' solution for for a number of years
for general data storage and management problems that many application developers face, we have noted a recent rise
in the use of non-relational data management tools~\cite{padhy2011rdbms}. Most of these tools have traditionally fallen into the domain of
big data analytics, with platforms such as the Hadoop ecosystem~\footnote{https://hadoop.apache.org/} being notably popular.
However, outside of the domain of big data, looking more at general purpose data storage and management, what are now
commonly referred to as ``NoSQL'' solutions are proving to be a popular solution.

NoSQL databases refer to those databases that are not built on top of the relational algebraic concepts, as laid out by
Codd in 1970~\cite{codd1970relational}, unlike the more commonly used RDBMS technologies, such as
MySQL~\footnote{https://www.mysql.com/}. Being free of the strictness the relational model enforces on its data allows
NoSQL databases to focus less on the overall structure of data, and more on factors such as scalability and
performance~\cite{leavitt2010will}.

While the relational model is a good fit for many data problems, its strictness in terms of flexibility of managing data
eventually led to the introduction of the NoSQL model. The following characteristics can be given as a starting point
for NoSQL databases in comparison to relational databases~\cite{indrawan2012database}:

\begin{itemize}
  \item \textbf{Unstructured data support:} While the relational model would often force data to be stored in tabular
  formats, the NoSQL model does not force any kind of data schema.
  \item \textbf{Designed with distributed processing and horizontal scalability in mind:} Given the commoditisation of
  computer hardware in the last decade, support for horizontal scaling and processing among clusters is an important
  factor for adoption.
  \item \textbf{Less strict adherence to ACID principles:} While the relational model attempted to very much adhere to
  the transactional principles of data atomicity, consistency, isolation, and durability (ACID), this very much impacts
  performance in terms of distributed computing. Relaxing the strictness of adherence to these principles, allow many
  NoSQL databases to make the trade-off for higher performance.
\end{itemize}

Of course, these differences between the NoSQL model and relational model vary between each individual database
technology's design, and trade-offs are often made depending on the goals and aims for that given database.

In this paper, we will look at the use of MongoDB~\footnote{https://www.mongodb.org/}, a popular NoSQL database solution,
as a solution for a case study based upon a railway data problem using data from Monash University's Institute of Railway
Technology (IRT). An overview of the case study in question will be given in~\sectref{sec:case_study_overview}. A small
overview of the MongoDB database will be given in~\sectref{sec:mongo_overview}. Implementation and
evaluation details will be given in~\sectref{sec:implementation} and~\sectref{sec:evaluation_and_discussion}, respectively,
before concluding in~\sectref{sec:conclusion}.

% section introduction (end)

\newpage

\section{Case Study Overview} % (fold)
\label{sec:case_study_overview}

The case study that is being looked at involves a project that has been worked at at the Monash University Institute of
Railway technology (IRT). The project involves trains the operate in the Pilbara region of Western Australia, taking
ore and minerals from loading points at mines to a specified unloading port. The data that is being dealt with comes from
numerous categories of sensors being placed on certain specialised train cars to record data monitoring track and train
car conditions~\cite{darby2003development,darby2005track}.

Data currently gets unloaded at sent back in large batches to remote servers once the car completes a trip and pulls back
into port~\cite{thomas2012taking}. The project currently makes use of a RDBMS solution to manage data storage, however
this solution is faced with many problems that currently require painful work-arounds. The most obvious of which involves
unreliable data being received from sensors. For example, in the case of a damaged or failed sensor, reliable data cannot
be guaranteed to be returned from such a sensor. Hence, data received by the remote server is often inconsistently
structured, and thus the data has to go through a series of preprocessing work arounds to fit in within the strict
schema that the RDBMS expects. As the only guaranteed consistency in the data that gets received is the timestamp and
geocoordinates, the strictly structured relational model is not an appropriate solution.

The Monash IRT team are currently investigating further solutions in the NoSQL big data space, where they intend to
replace the current system with an appropriately tested solution. While the scope of that project is much larger than
what will be covered in this paper, we will look at the differences in what is possible when attempting to use a NoSQL
database for the data storage and management.

For this paper, we will be looking at the possibility of using MongoDB, a popular NoSQL database that does away with
the concept of the schemas and tables so commonly found in databases following the relational model. This is done with
the expectation of better handling of the given inconsistent data.

% section case_study_overview (end)

\newpage

\section{MongoDB Overview} % (fold)
\label{sec:mongo_overview}

% section mongo_overview (end)

\newpage

\section{Implementation} % (fold)
\label{sec:implementation}

% section implementation (end)

\newpage

\section{Evaluation and Discussion} % (fold)
\label{sec:evaluation_and_discussion}

% section evaluation_and_discussion (end)

\newpage

\section{Conclusion} % (fold)
\label{sec:conclusion}

% section conclusion (end)

\newpage                            % Make references section

\bibliographystyle{acm}
\bibliography{biblio}

\end{document}